---
title: "Homework 5"
date: "Due 2018-10-11 at 8pm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

+ Follow [this link](https://classroom.github.com/a/vbOtC-cU) to create your 
repository for homework 5.

+ Once your repository is created, create a new RStudio project based on that
repository.

## Mice

The purpose of this problem is to build intuition for the bootstrap procedure.
With that in mind, consider the very small data set of the weights of three mice
that came from the same litter. The weights are 82, 107 and 93 grams. Please
write your responses in comments and your code in an R script named `mice.R` or `mice.Rmd`.

```{r, eval = FALSE}
weights <- c(82, 107, 93)
```

+ What is the mean weight of the three mice?
+ How many possible bootstrap samples are there from this sample of three mice?
+ List all possible bootstrap samples and compute the mean of each.
+ Compute the mean of the bootstrap means. How does this compare with the mean from part (a)
+ Calculate a percentile-based bootstrap confidence interval with level of confidence at least 95%. What information does this interval provide? How confident are you in reporting this interval? Explain.
 
## Trees
 
In the trees dataset found in R, there are three variables "Girth" "Height"
and "Volume". There is interest in using "Girth" and "Height" (two noninvasive
measures) to predict "Volume". Carrying out this prediction can be done using a
regression model with two explanatory variables ("Height" and "Girth"). This
model can very easily be fit in R using the following two lines of code.

```{r, eval = FALSE}
data(trees)
lm(Volume ~ Height + Girth, data=trees)
```

Principal interest lies in estimating the regression coefficient of "Height"
divided by the regression coefficient for "Girth". Obtaining a estimate of
the ratio is easy --- just compute the ratio of the two slope estimates ---
but assessing uncertainty associated with the point estimate can be
difficult theoretically (although methods do exist). 

+ For this problem construct an 95% confidence interval using the bootstrap 
procedure. Use 5,000 bootstrap samples. Compute your estimate and confidence 
bounds in an R script named `trees.R` or `trees.Rmd`.

## Showers

You think that your roommate takes showers whose duration is different from
yours. Your roommate disagrees but, instead of arguing, you both decide to keep
track of the time it takes both of you to shower for fourteen days and then to
employ skills from an introductory statistics class (e.g., STAT 121) to create a
confidence interval on the population mean showering time of your roommate minus
the population mean showering time of you. The data and code to compute a 90%
confidence interval are below.

```{r, eval = FALSE}
roommate <- c(6,6.25,5.75,6.5,6,15,35,20,4,5.25,5.75)

you <- c(3.5,5.5,4,4,4,6,11,12,3,4,4.25,7,3.25)

t.test(roommate, you, conf.level = 0.9)
```

Notice that the confidence interval contains 0 and, therefore, your roommate
concludes that there is no significant difference in the time it takes you
both to take a shower. But it might be inappropriate to use this confidence
interval on the difference in means because the data are not normally
distributed and the sample sizes are small. 

+ Use the bootstrap procedure to compute a 90% confidence interval on the difference in these means. Based on
this bootstrap confidence interval, what conclusion do you make at the 0.10
level of significance? Please write your responses in comments and your code
in an R script named `shower.R` or `shower.Rmd`.

#### **Make sure you commit your changes to git and push to GitHub**
