---
title: "Coin Flip Simulation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```


### Single coin
Randomly sample one value from the set [0, 1] where the probability of 0 is 3/4
and the probability of 1 is 1/4.

```{r}
sample(c(0, 1), 1, prob = c(3/4, 1/4))
```

Simulate the process of tossing a bent coin one time, where the probability of a
head is p = 1/4

```{r}
sample(c(0, 1), 1, prob = c(3/4, 1/4))
```

### Multiple coins
Simulate the process of tossing a bent coin 60 times, where the probability of a
head is p = 1/4.

```{r}
sample(c(0, 1), 60, prob = c(3/4, 1/4), replace = TRUE)
sample(c(0, 1), 60, prob = c(3/4, 1/4), replace = TRUE)
```

Notice that the results above are different - why?

You can use `set.seed()` to ensure results are identical when random elements are
involved.

```{r}
set.seed(35487)
sample(c(0, 1), 60, prob = c(3/4, 1/4), replace = TRUE)

set.seed(35487)
sample(c(0, 1), 60, prob = c(3/4, 1/4), replace = TRUE)


```

### Simulation study
Now that we have the building blocks, we'll simulate this process multiple times
to provide a reliable estimate of the average number of heads out of 60 tosses.
There are multiple approaches that can be used here, and we'll look at each one.

In order to build these simulation examples, let's setup a few things first. We'll
define the number of simulations we run as `n_sims` and we'll also create a 
function for a single "run" of the simulation.

```{r}
n_sims <- 1000

flip_coins <- function(n_flips = 60) {
  sample(c(0, 1), n_flips, prob = c(3, 1), replace = TRUE)
}
```

```{r}
flip_coins()
```


#### For Loop
For loops are a staple programming concept. However, given R's ability to vectorize
operations and other R-centric functions (the `apply` family), there are often
better ways to iterate through objects.

```{r}
heads <- numeric(n_sims)
for (i in seq_along(heads)) {
  heads[i] <- sum(flip_coins())
}

heads
```

Now that we have the number of heads that appeared in 1000 simulations of flipping
this coin 60 times, we can compute the average number of heads that appear in this
scenario.

```{r}
mean(heads)
```


#### Sapply function

```{r}
heads <- sapply(1:n_sims, function(x) sum(flip_coins()))
```

#### Replicate function
`replicate` is part of the `apply` family and it lends itself well to simulation
studies.

```{r}
heads <- replicate(n = n_sims, sum(flip_coins()))
```

#### Tidyverse
The [tidyverse](https://www.tidyverse.org/) " is an opinionated collection of R 
packages designed for data science. All packages share an underlying design 
philosophy, grammar, and data structures."

The `purrr` package is part of the `tidyverse` and provides an alternative to 
the `apply` family of functions.

```{r}
# install.packages("tidyverse")
library(tidyverse)

rerun(n_sims, flip_coins()) %>% 
  map_dbl(sum) %>% 
  mean()
```


### Calculations
So far, we've used our simulated data to determine the average number of heads
in a set of 60 unfair coin flips. We can also use this data to calculate other
metrics.

Determine the variance of the number of heads in 60 coin flips.

```{r}
var(heads)
```

We can also find the probability that the number of heads in 60 tosses is greater
than 20.
```{r}
mean(heads > 20)
```

Finally, we can ind the 20th percentile of the number of heads in 60 tosses.
```{r}
quantile(heads, 0.25)
```

