---
title: "Central Limit Theorem"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```


The [Central Limit Theorem](https://en.wikipedia.org/wiki/Central_limit_theorem)
is "a key concept in probability theory because it implies that probabilistic and 
statistical methods that work for normal distributions can be applicable to many
problems involving other types of distributions."

Consider a right skewed distribution (e.g Binomial(10, 0.1))

```{r}
n_trials <- 10
pr_success <- 0.1

x <- 0:n_trials
y <- dbinom(x, n_trials, pr_success)
```

Now let's visualize our results.

```{r}
plot(NA, xlim = c(0, n_trials), ylim = c(0, 0.4))
invisible(sapply(1:length(x), function(i) lines(c(x[i], x[i]), c(0, y[i]))))
```

```{r}
library(tidyverse)

tibble(x, y) %>% 
  ggplot(aes(x = x, y = y)) +
  ggalt::geom_lollipop()
```

Now, let's construct sampling distributions of the mean when samples of size 40
are drawn from this population.

```{r}
sample_size <- 40
n_draws <- 1000
```

#### For Loop

```{r}
sample_means <- numeric(n_draws)
for (i in seq_along(sample_means)) {
  sample_means[i] <- mean(rbinom(sample_size, n_trials, pr_success))
}
```

Now, let's investigate the resulting sample means.

```{r}
hist(sample_means)
```

```{r}
mean(sample_means)
```

```{r}
var(sample_means)
```

#### sapply

```{r}
sample_means <- sapply(1:n_draws, function(x){
 mean(rbinom(sample_size, n_trials, pr_success))
})
```


#### Tidyverse

```{r}
library(tidyverse)

sample_means <- rerun(n_draws, rbinom(sample_size, n_trials, pr_success)) %>% 
  map_dbl(mean)
```


---

Write a function whose input is `n`, `n_trials`, `pr_success`, and `n_draws`
which returns the mean of `n_draws` samples of size `n` from a Binomial(`n_trials`,
`pr_success`) distribution. Let `n_draws` be 1000 by default.

```{r}
sample_dist <- function(n, n_trials, pr_success, n_draws = 1000) {
  replicate(n_draws, mean(rbinom(n, n_trials, pr_success)))
}
```

```{r}
demo_sample_dist <- function(n) {
  hist(sample_dist(sample_size, n_trials, pr_success), breaks = 30, freq = FALSE)
  x <- seq(0, 3, length = 1000)
  pop_mean <- n_trials * pr_success
  pop_variance <- pop_mean * (1 - pr_success) / n
  lines(x, dnorm(x, pop_mean, sqrt(pop_variance)), col = "red")
}
```

```{r}
demo_sample_dist(40)
```

This is a manifestation of the Central Limit Theorem.
