---
title: "Mean"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

Consider the following data:

```{r}
x <-
  c(2.11,
    0.87,
    0.95,
    1.51,
    0.28,
    1.45,
    0.18,
    0.52,
    -1.16,
    -0.56,
    0.13,
    0.65,
    -0.01,
    0.95,
    -1.95)
```

Given this data, we have the following assumptions:

* Independence
* Known standard deviation
* Normally distributed population

In this case, our known population standard deviation is 1.

```{r}
sigma <- 1
```

We want to conduct the following hypothesis test:

$$
H_0: \mu = 0
$$

$$
H_a: \mu > 0
$$

$$
\alpha = 0.05
$$

First, we'll create a function using the following formula to give us our z test
statistic.

$$
z = \frac{\bar{x} - \mu}{\sigma / \sqrt{n}}
$$

```{r}
alpha <- 0.05
z_test_stat <- function(x, mu, sigma) {
  x_bar <- mean(x)
  n <- length(x)
  (x_bar - mu) / (sigma / sqrt(n))
}
```

Now we'll conduct the hypothesis test.

```{r}
z <- z_test_stat(x, 0, 1)
z
```

```{r}
p_val <- pnorm(z, mean = 0, sd = 1, lower.tail = FALSE)
p_val
```

What is the critical value (how big does `z` need to be)?
```{r}
critical_val <- qnorm(alpha, mean = 0, sd = 1, lower.tail = FALSE)
critical_val
```

Conclusion: Fail to reject $H_0$ since p-value is greater than $\alpha = 0.05$.

This approach requires theory to know that the sampling distribution of the Z
statistic has a standard normal distribution.

We could instead use a simulation study to find the sampling distribution of Z
and compute p-values, power, etc.

```{r}
simulate <- function(n, mu, sd, n_reps = 1000) {
  replicate(n_reps, z_test_stat(rnorm(n, mu, sd), mu, sd))
}
```

```{r}
test_stat_under_h0 <- simulate(length(x), 0, 1)
test_stat_under_h0
```

```{r}
hist(test_stat_under_h0)
```

Now that we have a distribution of test statistics, we can once again perform
our hypothesis test, this time leaning on our simulation instead of leaning on 
theory.

```{r}
p_val <- mean(test_stat_under_h0 >= z)
```

```{r}
ci <- function(x) {
  p <- mean(x)
  ci <- p * qnorm(0.975) * sqrt(p * (1 - p) / length(x))
  c(lower = p - ci,
    p_value = p,
    upper = p + ci)
}

ci(test_stat_under_h0 >= z)
```





