---
title: "Z Test on Proportions"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

Consider a two-sided z-test on a proportion $p$ based on $x$ successes in $n$
trials. The null hypothesis is that $p = p_0$ and the alternative hypothesis
is that $p != p_0$. The test says to reject the null hypothesis whenever the test
statistic $z = (\hat{p} - p_0) / \sqrt{(p_0 * (1 - p_0) / n)}$ is greater than
or equal to 1.96 in absolute value ($\hat{p} = x/n$). This test is reported to
have an $\alpha = 0.05$ level of significance, but the actual probability of a 
Type I error (rejecting the null hypothesis when in fact it is true) may be 
different from $\alpha$ because:

1. The test statistic is only approximately normally distributed
2. The actual distribution of the test statistic is discrete

Compute the actual probability of a Type I error for the procedure described for
all combinations of $p = [0.1, 0.3, 0.5]$ and $n = [20, 30, ..., 100]$. Present
the results in a 9 x 3 matrix, with the sample $n$ going down the rows and the 
values of $p$ across the columns.

Assess the Monte Carlo error with a 95% confidence interval for each combination
of $p$ and $n$. For which combinations does the lower bound on the confidence
interval exceed $\alpha = 0.05$?


```{r}
simulate <- function(n, p, alpha = 0.05, n_reps = 100000) {
  x <- rbinom(n_reps, n, p)
  p_hat <- x / n
  z <- (p_hat - p) / sqrt(p * (1 - p)/n)
  # What proportion of times did we commit a Type I error at the given alpha level
  mean(abs(z) >= qnorm(1 - alpha/2))
}
```

```{r}
simulate(30, .3)
```

```{r}
p_seq <- c(0.1, 0.3, 0.5)
n_seq <- seq(20, 100, by = 10)

est <- matrix(NA, nrow = length(n_seq), ncol = length(p_seq))
dimnames(est) <- list(sprintf("n=%d", n_seq),
                      sprintf("p=%.1f", p_seq))
est
```

```{r}
n_reps <- 100000
# Iterate over rows
for (i in 1:length(n_seq)) {
  # Iterate over columns
  for (j in 1:length(p_seq)) {
    est[i,j] <- simulate(n_seq[i], p_seq[j], n_reps = n_reps)
  }
}

est
```


```{r tidy, message=FALSE}
library(tidyverse)
results <- expand.grid(p = p_seq, n = n_seq) %>% 
  mutate(value = map2_dbl(n, p, simulate, n_reps = n_reps)) %>% 
  spread(key = p, value = value)

results
```

```{r ci}
ci <- qnorm(0.975) * sqrt(est * (1 - est)/n_reps)
lower <- est - ci

alpha <- 0.05
lower > alpha
```


