---
title: "M&M Study"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

My daughter is convinced she can tell the difference between M&M colors by their
flavor. I'm not buying it. Since I'm a statistician, we devise the following
two experiments in an attempt to investigate her claim.

### Experiment 1
There are 6 M&M colors. In this experiment, I select one M&M of each color and 
give it to my daughter while she is blindfolded. She guesses the color of each
M&M as she tastes it. We agree that after she guesses a color, she will not guess
it again in that round. We perform this exercise for three rounds, so a total of 
18 M&Ms are consumed. At the end, we count the number of colors she correctly 
determined out of 18.

### Experiment 2
In this version, I randomly draw 18 M&Ms from the bag, without regard to color, 
and once again provide them to my blindfolded daughter. She makes her best guess
at the color of each M&M and at the conclusion we count the number of correctly
determined M&Ms out of 18.

Given these experiments, we want to answer the following questions:

1. When using $\alpha = 0.5$ level of significance, what is the critical value 
for testing the null hypothesis that she is merely guessing (equal likelihood of
selecting any given color)? What is the size of the test?

2. Suppose my daughter can correctly distinguish blue M&Ms from the others, but
if the M&M is not blue, she's merely guessing. What is the statistical power of 
the testing procedure for this situation? Given the answer, which experiment
would be preferred by a statistician (assuming both are equally convenient)?

3. Suppose my daughter correctly identified 7 M&Ms. Is this sufficient evidence 
against the null hypothesis that she is merely guessing?

```{r}
exp_1 <- function(h0 = TRUE) {
  max <- ifelse(h0, 6, 5)
  truth <- sample(1:max)
  guess <- sample(1:max)
  # How many were correctly guessed?
  sum(guess == truth) + !h0
}
```

```{r}
sim_1 <- function(h0 = TRUE, n_rounds = 3, n_reps = 10000) {
  replicate(n_reps, sum(replicate(n_rounds, exp_1(h0))))
}
```

```{r}
sim_2 <- function(h0 = TRUE, n_samples = 18, n_reps = 10000) {
  if (h0) {
    rbinom(n_reps, n_samples, p = 1/6)
  } else {
    blues <- rbinom(n_reps, n_samples, p = 1/6)
    others <- rbinom(n_reps, n_samples - blues, p = 1/5)
    blues + others
  }
}
```

```{r}
h0_dist_1 <- sim_1()
ha_dist_1 <- sim_1(FALSE)
```

```{r}
h0_dist_2 <- sim_2()
ha_dist_2 <- sim_2(FALSE)
```

Now that we have these distributions, we can answer the questions.

```{r}
alpha <- 0.05
# Critical value
crit_val_1 <- quantile(h0_dist_1, probs = 1 - alpha, type = 1) + 1
# Size (probability of Type I error (false rejection of null hypothesis))
mean(h0_dist_1 >= crit_val_1)
```

```{r}
# Critical value
crit_val_2 <- quantile(h0_dist_2, probs = 1 - alpha, type = 1) + 1
# Size (probability of Type I error (false rejection of null hypothesis))
mean(h0_dist_2 >= crit_val_2)
```

Statistical power is "the probability the test correctly rejects $H_0$ for a given
$H_A$.

```{r}
prop_ci <- function(x) {
  est <- mean(x)
  ci <- qnorm(0.975) * sqrt(est * (1 - est) / length(x))
  c(
    lower = est - ci,
    estimate = est,
    upper = est + ci
  )
}
```


```{r}
power_1 <- prop_ci(ha_dist_1 >= crit_val_1)
power_1
```

```{r}
power_2 <- prop_ci(ha_dist_2 >= crit_val_2)
power_2
```

```{r}
# Number of M&Ms correctly identified
observed_stat <- 7

p_val_1 <- prop_ci(h0_dist_1 >= observed_stat)
p_val_1
```

```{r}
p_val_2 <- prop_ci(h0_dist_2 >= observed_stat)
p_val_2
```
