---
title: "Bootstrap Confidence Interval on a mean"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

Bootstrapping is a method that relies on random sampling with replacement. This
method allows us to build measures of accuracy or perform inference in situations
where other methods prove difficult.

Let's illustrate this via a simple example. Suppose we have 15 observations drawn
from a random normal distribution.

```{r}
set.seed(35487)
n <- 15
x <- rnorm(n)
```

We can plot the density of this sample to confirm that it's approximately normal.

```{r}
plot(density(x))
```

We can use a t-test to provide a confidence interval based on the normality assumption:

```{r}
t.test(x)
```

Now, we could also create a confidence interval by building a series of bootstrapped
samples from our given sample of 15 observations. This involves sampling 15
observations from our original sample **with replacement** and calculating the mean
value of these bootstrapped samples. It's important to note that since we are
sampling with replacement, there are $15^{15}$, or $`r 15^15`$ possible bootstrap
samples. While we could *theoretically* include each of those samples, that would
take an exceptional amount of time and compute power. Instead, we'll once again
rely on the central limit theorem and assume that if we provide a large enough 
collection of bootstrap samples, we will be ok.

```{r}
n_samples <- 10000
bootstrap_means <- replicate(n_samples, mean(sample(x, replace = TRUE)))
```

Now that we have these 10,000 sample means constructed from our original sample,
we can explore the density and then use our collection of means to provide an
estimated confidence interval.

```{r}
plot(density(bootstrap_means))
```

```{r}
ci <- quantile(bootstrap_means, probs = c(0.025, 0.975))
ci
```

Finally, we can visualize this confidence interval along with the density we 
previously plotted.

```{r}
plot(density(bootstrap_means))
abline(v = ci, col = "red")
```

