---
title: "Memory Loss and Marijuana Use"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

## Two sample t-test of difference of means 
As part of a research project on the effects that marijuana smoking has on 
short-term memory, two groups of ten subjects were tested. One group, the 
"nonsmokers", claimed not to smoke marijuana. A second group, the "smokers," 
claimed to smoke marijuana regularly. The following are the number items remembered 
successfully on the digit span task.

```{r}
nonsmokers <- c(18, 22, 21, 17, 20, 17, 23, 20, 22, 21)
smokers <- c(16, 20, 14, 21, 20, 18, 13, 15, 17, 21)
n1 <- length(nonsmokers)
n2 <- length(smokers)
mean(nonsmokers)
mean(smokers)
```

Two-sample t-test:  $H_0: \mu_1 = \mu_2$
Assumptions:
1. Independent observations from both populations
2. Observations from both groups are drawn from normally distributed populations
3. The two populations have the same variance (if not using Welch's test)

```{r}
tt <- t.test(nonsmokers, smokers, var.equal = TRUE)
tt
```

Welch's t-test

```{r}
ttw <- t.test(nonsmokers, smokers, var.equal = FALSE)
ttw
```

Assumptions from bootstrap procedure
1. Independent observations from both populations

```{r}
observed_difference <- mean(nonsmokers) - mean(smokers)
observed_difference
```

Note that what we're interested in here is the value of this `observed_difference`
as a way of testing a difference in means. Now, we can use bootstrap sampling of
both samples to build a distribution around the difference in means.

```{r}
n_samples <- 10000
sample_differences <- replicate(n_samples, mean(sample(nonsmokers, replace = TRUE)) - mean(sample(smokers, replace = TRUE)))
```

Let's investigate the distribution of `sample_differences`.

```{r}
plot(density(sample_differences))
```

Using this distribution, we can build a confidence interval around the difference
in sample means.

```{r}
ci <- quantile(sample_differences, probs = c(0.025, 0.975))
ci
```

```{r}
plot(density(sample_differences))
abline(v = ci, col = "red")
```

Using this method, we come to the same conclusion that we did using the t-test, 
namely that we can reject the null hypothesis.
