---
title: "Bootstrap Confidence Interval"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

One of the conveniences of the bootstrap method is that it allows us to create a
confidence interval around arbitrary metrics.

Consider the `faithful` dataset in R that contains information about Old Faithful
eruptions in Yellowstone. The `waiting` column contains data on the time between
eruptions.

```{r}
hist(faithful$waiting)
```

We want to estimate the 25th quantile for this data. We can calculate that with
the following:

```{r}
quantile(faithful$waiting, 0.25)
```

Now, how do we measure our uncertainty in this measurement? If we don't have a 
handy formula, can we still measure uncertainty? Bootstrap to the rescue!

+ Create `n` bootstrap samples of `faithful$waiting`
+ For each bootstrap sample, calculate the 25th quantile value
+ Given this distribution of `n` 25th quantile values, build a confidence interval
on the 25th quantile.

```{r}
n_samples <- 10000
quantile_samples <- replicate(n_samples, quantile(sample(faithful$waiting, replace = TRUE), .25))
```

```{r}
quantile(quantile_samples, c(0.025, 0.975))
```

Now, we can expand what we've done to create a generic solution that provides a
bootstrap confidence interval for an arbitrary metric.

```{r}
bootstrap_ci <- function(x, f, alpha = 0.05, n_samples = 5000) {
  bs_out <- replicate(n_samples, f(sample(x, replace = TRUE)))
  quantile(bs_out, c(alpha/2, 1 - alpha/2))
}
```

```{r}
bootstrap_ci(faithful$waiting, function(x) quantile(x, .25), alpha = .05)
```
