---
title: "Bootstrap Coverage"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, message = FALSE)
```

Using the idea of coverage that was introduced with simulation studies, we can 
assess the coverage of bootstrapped confidence intervals.

Consider a random sample from a standard normal distribution.

```{r}
n <- 50
x <- rnorm(n)
```

Given this, we want to estimate the 25th quantile value.
```{r}
quantile(x, 0.25)
```

We can build a confidence interval around this estimate using the bootstrap method.
```{r}
bootstrap_ci <- function(x, f, alpha = 0.05, n_samples = 1000) {
  bs_out <- replicate(n_samples, f(sample(x, replace = TRUE)))
  quantile(bs_out, c(alpha/2, 1 - alpha/2))
}
```

```{r}
bootstrap_ci(x, function(x) quantile(x, 0.25))
```

We also can calculate the actual value of the 25th quantile of the standard normal
distribution.
```{r}
truth <- qnorm(0.25)
truth
```

Given that we know the actual value, we can use a simulation study to assess the
coverage of our bootstrap confidence interval.

```{r}
library(tictoc)
n_reps <- 1000
```

```{r}
tictoc::tic()
bs_ci_values <- replicate(n_reps, bootstrap_ci(rnorm(n), function(x) quantile(x, 0.25)))
tictoc::toc()
```

```{r}
dim(bs_ci_values)
```

Notice that the above code takes a *long* time to run. Why?

We can speed things up by running our code in parallel! First, we need to install
a couple of useful packages, `future.apply` and `furrr`. `future.apply` provides
parallelized versions of the R `apply` functions while `furrr` provides parallelized
versions of the `purrr` map functions.

```{r, eval=FALSE}
install.packages(c("future.apply", "furrr"))
```

```{r}
library(future.apply)
library(furrr)
```

Now, we need to setup our computer to take advantage of the cores we have available.

```{r}
availableCores()
plan("multiprocess")
```

# Run the same code as above, but this time in parallel using `future_replicate()`.

```{r}
tic()
bs_ci_values <- future_replicate(n_reps, bootstrap_ci(rnorm(n), function(x) quantile(x, 0.25)))
toc()
```

```{r}
dim(bs_ci_values)
```


```{r}
tic()
bs_ci_list <- future_map(1:n_reps, 
                         function(x) bootstrap_ci(rnorm(n), function(x) quantile(x, 0.25)))
toc()
```

```{r}
length(bs_ci_list)
```

Now that we have these confidence interval values, we can estimate coverage by
determining what proportion of intervals contain the true value.

```{r}
coverage <- mean(purrr::map_lgl(bs_ci_list, ~.[1] <= truth & .[2] >= truth))
coverage
```

```{r}
coverage_ci <- coverage + c(-1, 1)*qnorm(0.975)*sqrt(coverage*(1 - coverage)/n_reps)
c(lower = coverage_ci[1],
  coverage = coverage,
  upper = coverage_ci[2])
```
