---
title: "High School and College GPA"
subtitle: "Is there a relationship?"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

Consider the following data that reprents high school and college GPAs.

```{r}
gpa <- data.frame(
  hsgpa = c(3.1, 2.3, 3, 2.5, 3.9, 2.25, 4, 3.8, 3.7, 3.3, 3.8, 3.6, 3.6, 3.3, 3.4, 3.4, 3, 2.9, 3.1, 3.5, 3.3, 2.6, 3, 3.6, 2),
cgpa = c(1.17, 2.24, 2.42, 1.8, 3.16, 1.18, 2.71, 3.52, 2.92, 3.4, 2.02,  3.38, 2.69, 3.04, 2.35, 2, 2.32, 2.48, 2.37, 3.14, 1.88, 2.17, 2.3, 2.18, 0)
)
```

```{r}
plot(gpa)
```

Given this data, we want to know if there is a significant correlation between
high school and college GPA.

+ Calculate the observed correlation (`cor()`)
+ Use a permutation test to build a distribution of values under the null hypothesis
(there is no correlation)
+ Use this distribution to determine the p value of the observed correlation
+ Build a confidence interval around the p value

```{r}
observed_value <- cor(gpa$hsgpa, gpa$cgpa)
observed_value
```

```{r}
n_permutations <- 10000
results <- replicate(n_permutations, {
  new_gpa <- data.frame(hsgpa = gpa$hsgpa, cgpa = sample(gpa$cgpa))
  cor(new_gpa$hsgpa, new_gpa$cgpa)
})
```

```{r}
plot(density(results))
abline(v = observed_value, col = "red")
```

```{r}
p_value <- mean(results >= observed_value)
p_value
```

```{r}
ci <- p_value + c(-1, 1) * qnorm(.975) * sqrt(p_value * (1 - p_value) / n_permutations)
c(lower = ci[1],
  p_value = p_value,
  upper = ci[2])
```

