---
title: "Old Faithful - Time Series Autocorrelation"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

In this example we're going to use the built in `faithful` dataset which contains
two columns: `eruptions` and `waiting`. We want to determine of the waiting times
between Old Faithful eruptions are autocorrelated.

$$
H_0: \text{there is no correlation} \\
H_a: \text{there is a correlation}
$$

First, let's look at the data for Old Faithful waiting times.

```{r}
waiting <- faithful$waiting
plot(waiting, type = "b")
```

Now, to estimate autocorrelation, we can do the following:
```{r}
observed_cor <- cor(waiting[-length(waiting)], waiting[-1])
observed_cor
```

Now, in order to determine if this value is significant, we need some way to 
determine a p value. Once again, we can rely on permutation testing to build a 
distribution of values under the null hypothesis and then compare our observed
value to that distribution.

```{r}
n_permutations <- 10000
results <- replicate(n_permutations, {
  new_waiting <- sample(waiting)
  cor(new_waiting[-length(new_waiting)], new_waiting[-1])
})
```

Now that we have `results`, which contains 10000 autocorrelation values under the
null hypothesis, we can examine its distribution.

```{r}
plot(density(results))
```

Now, let's look at our `observed_cor` value in light of these results.

```{r}
plot(density(results))
abline(v = observed_cor, col = "red")
```

Given our `observed_cor` and the values contained in `results`, we can caluclate
a p value and associated confidence interval as follows:
```{r}
p_value <- mean(abs(results) >= abs(observed_cor))
p_value
```

```{r}
ci <- p_value + c(-1, 1) * qnorm(.975) * sqrt(p_value * (1 - p_value) / n_permutations)
c(lower = ci[1],
  p_value = p_value,
  upper = ci[2])
```

